Type Refinements for the Working Class

### 2015-12-21 paper.tex
update a bunch of instances of Sets^Sctx

  So, $\Refines{\Upsilon}{\phi}{\tau}$ obtains just when we have
  $\Member{\phi}{\PERs{\Exprs_\tau}(\Upsilon)}$. The benefit of explaining such
  judgments as copresheaves is that we do not need to deal with
  the complexities of quantifying over renamings, since this is implicit in the
  definition of the exponential of presheaves, where
  $\Define{\Coyoneda{\Upsilon}}{\Upsilon\inj-}$ is the co-yoneda embedding:
  \begin{align*}
    \ADefine{
      B^A(\Upsilon)
    }{
      \Hom{\Sets^\SCtx}{
        \Coyoneda{\Upsilon}\times A
      }{
        B
      }
    }\\
    &\cong
    \OMode{
      \int_{\Upsilon'}
      (\Upsilon\inj\Upsilon')
      \implies
      {B(\Upsilon')}^{A(\Upsilon')}
    }
  \end{align*}

@@

\begin{notation*}
  For a family of copresheaves $\HypJ{\Of{X_i}{\Sets^\SCtx}}{\Member{i}{I}}$, we will
  use $X^{[\vec{i}]}$ as a shorthand for the following iterated product, as
  in \cite{sterling-morrison:2015}:

@@

Fix a sort-indexed family of binary relations on closed expressions
$\Of{R_\tau}{\Hom{\Sets^{\SCtx}}{\mathbb{1}}{\Pow{\Exprs_\tau^2}}}$; we will
also write the relation as a judgment scheme
$\SBinRel{R_\tau}{\Upsilon}{M}{N}$. First of all, we can lift such a relation
to operate on open terms by quantifying over substitutions, as follows:
\[
  \infer{
    \OSBinRel{R_\tau}{\Upsilon}{\vec{x}:\vec{\sigma}}{M}{N}
  }{
    \forall
      \begin{tikzcd}[cramped, ampersand replacement = \&]
         \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
      \end{tikzcd}.
    \quad
    \forall \Of{\vec{X}}{\Exprs^{[\vec{\sigma}]}(\Upsilon')}.
    \quad
    \SBinRel{R_\tau}{\Upsilon'}{
      \Subst{\vec{X}}{\vec{x}}{\Rename{\rho}{M}}
    }{
      \Subst{\vec{X}}{\vec{x}}{\Rename{\rho}{M}}
    }
  }
\]

Next, we can always extend $R_\tau$ to a new
family of valenced-indexed relations
$\Of{R_v}{\Hom{\Sets^{\SCtx}}{\mathbb{1}}{\Pow{\BTms_v^2}}}$ for
$\Match{v}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$, defined pointwise as
follows:
\[
  \infer{
    \SBinRel{R_{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}}{\Upsilon}{
      \MkBTm{\vec{u}}{\vec{x}}{M}
    }{
      \MkBTm{\vec{v}}{\vec{y}}{N}
    }
  }{
    \begin{array}{l}
      \forall\IMode{\vec{w}}\mathrel{\#}\IMode{\Dom{\Upsilon}\cup\vec{u}\cup\vec{v}}.\quad
      \forall\IMode{\vec{z}}\mathrel{\#}\IMode{\vec{x}\cup\vec{y}}.\\
      \OSBinRel{R_\tau}{\Upsilon,\vec{w}:\vec{\sigma}}{\vec{z}:\vec{\tau}}{
        \Subst{\vec{z}}{\vec{x}}{
          \Rename{(\vec{u}\mapsto\vec{w})}{M}
        }
      }{
        \Subst{\vec{z}}{\vec{y}}{
          \Rename{(\vec{v}\mapsto\vec{w})}{M}
        }
      }
    \end{array}
  }
\]

@@

\begin{definition}[Computational Approximation]
  Because $[-_\tau]$ is monotonic, it has a greatest fixed point, which we will
  call \emph{computational approximation},
  $\Of{\preccurlyeq_\tau}{\Hom{\Sets^\SCtx}{\mathbb{1}}{\Pow{\Exprs_\tau^2}}}$,
  written pointwise as $\SBinRel{\preccurlyeq_\tau}{\Upsilon}{M}{N}$.
\end{definition}

@@

A sort-indexed family of binary relations
$\HypJ{\Of{R_\tau}{\Hom{\Sets^\SCtx}{\mathbb{1}}{\Pow{\Exprs_\tau^2}}}}{\Member{\tau}{\Sorts}}$
is a \emph{congruence} when for all operators
$\IsOperator{\Upsilon}{\vartheta}{\MkArity{\vec{v}}{\tau}}$, the following rule
obtains:
\[
  \infer{
    \SBinRel{R_\tau}{\Upsilon}{\vartheta(\vec{E})}{\vartheta(\vec{F})}
  }{
    \SBinRel{R_{\vec{v}}}{\Upsilon}{\vec{E}}{\vec{F}}
  }
  \tag{Congruence}
\]

@@

Now, let us return briefly to Martin-L\"of's notion of type, which we call a
\emph{value type}. The minimal data of such a type's definition consists in a
partial equivalence relation on values,
$\Of{\phi}{\Hom{\Sets^\SCtx}{\Coyoneda{\Upsilon}}{\PERs{\Values_\tau}}}$.  It
is easy to turn such a relation into a computational type
$\IsType{\Upsilon}{\VToCType(\phi)}{\tau}$, as follows, where $\lambda(\chi)$
is the exponential transpose of $\chi$:
\[
  \begin{tikzcd}[sep=huge]
    \IMode{\Coyoneda{\Upsilon}\times\Values_\tau^2}
      \arrow[r, "\IMode{\lambda(\phi)}"]
      \arrow[d, hook, swap, "\IMode{\Pair{\ArrId{}}{i}}"]
      &
    \IMode{\Omega}\\
    \IMode{\Coyoneda{\Upsilon}\times\Exprs_\tau^2}
    \arrow[ur, swap, dotted, "\OMode{\lambda(\VToCType(\phi))}"]
  \end{tikzcd}
\]
where $\VToCType(\phi)$ is the computational extension of $\phi$ in the following sense:
\begin{align*}
  \ADefine{
    \VToCType(\phi)(M,N)
  }{
    \forall M', N' \in \Values_\tau.\
    M\sim_\tau M' \land N\sim_\tau N' \implies \phi(M',N')
  }
\end{align*}

>>>>

  So, $\Refines{\Upsilon}{\phi}{\tau}$ obtains just when we have
  $\Member{\phi}{\PERs{\Exprs_\tau}(\Upsilon)}$. The benefit of explaining such
  judgments as copresheaves is that we do not need to deal with
  the complexities of quantifying over renamings, since this is implicit in the
  definition of the exponential of presheaves, where
  $\Define{\Coyoneda{\Upsilon}}{\Upsilon\inj-}$ is the co-yoneda embedding:
  \begin{align*}
    \ADefine{
      B^A(\Upsilon)
    }{
      \Hom{\SemU}{
        \Coyoneda{\Upsilon}\times A
      }{
        B
      }
    }\\
    &\cong
    \OMode{
      \int_{\Upsilon'}
      (\Upsilon\inj\Upsilon')
      \implies
      {B(\Upsilon')}^{A(\Upsilon')}
    }
  \end{align*}

@@

\begin{notation*}
  For a family of copresheaves $\HypJ{\Of{X_i}{\SemU}}{\Member{i}{I}}$, we will
  use $X^{[\vec{i}]}$ as a shorthand for the following iterated product, as
  in \cite{sterling-morrison:2015}:

@@

Fix a sort-indexed family of binary relations on closed expressions
$\Of{R_\tau}{\Hom{\SemU}{\mathbb{1}}{\Pow{\Exprs_\tau^2}}}$; we will
also write the relation as a judgment scheme
$\SBinRel{R_\tau}{\Upsilon}{M}{N}$. First of all, we can lift such a relation
to operate on open terms by quantifying over substitutions, as follows:
\[
  \infer{
    \OSBinRel{R_\tau}{\Upsilon}{\vec{x}:\vec{\sigma}}{M}{N}
  }{
    \forall
      \begin{tikzcd}[cramped, ampersand replacement = \&]
         \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
      \end{tikzcd}.
    \quad
    \forall \Of{\vec{X}}{\Exprs^{[\vec{\sigma}]}(\Upsilon')}.
    \quad
    \SBinRel{R_\tau}{\Upsilon'}{
      \Subst{\vec{X}}{\vec{x}}{\Rename{\rho}{M}}
    }{
      \Subst{\vec{X}}{\vec{x}}{\Rename{\rho}{M}}
    }
  }
\]

Next, we can always extend $R_\tau$ to a new
family of valenced-indexed relations
$\Of{R_v}{\Hom{\SemU}{\mathbb{1}}{\Pow{\BTms_v^2}}}$ for
$\Match{v}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$, defined pointwise as
follows:
\[
  \infer{
    \SBinRel{R_{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}}{\Upsilon}{
      \MkBTm{\vec{u}}{\vec{x}}{M}
    }{
      \MkBTm{\vec{v}}{\vec{y}}{N}
    }
  }{
    \begin{array}{l}
      \forall\IMode{\vec{w}}\mathrel{\#}\IMode{\Dom{\Upsilon}\cup\vec{u}\cup\vec{v}}.\quad
      \forall\IMode{\vec{z}}\mathrel{\#}\IMode{\vec{x}\cup\vec{y}}.\\
      \OSBinRel{R_\tau}{\Upsilon,\vec{w}:\vec{\sigma}}{\vec{z}:\vec{\tau}}{
        \Subst{\vec{z}}{\vec{x}}{
          \Rename{(\vec{u}\mapsto\vec{w})}{M}
        }
      }{
        \Subst{\vec{z}}{\vec{y}}{
          \Rename{(\vec{v}\mapsto\vec{w})}{M}
        }
      }
    \end{array}
  }
\]

@@

\begin{definition}[Computational Approximation]
  Because $[-_\tau]$ is monotonic, it has a greatest fixed point, which we will
  call \emph{computational approximation},
  $\Of{\preccurlyeq_\tau}{\Hom{\SemU}{\mathbb{1}}{\Pow{\Exprs_\tau^2}}}$,
  written pointwise as $\SBinRel{\preccurlyeq_\tau}{\Upsilon}{M}{N}$.
\end{definition}

@@

A sort-indexed family of binary relations
$\HypJ{\Of{R_\tau}{\Hom{\SemU}{\mathbb{1}}{\Pow{\Exprs_\tau^2}}}}{\Member{\tau}{\Sorts}}$
is a \emph{congruence} when for all operators
$\IsOperator{\Upsilon}{\vartheta}{\MkArity{\vec{v}}{\tau}}$, the following rule
obtains:
\[
  \infer{
    \SBinRel{R_\tau}{\Upsilon}{\vartheta(\vec{E})}{\vartheta(\vec{F})}
  }{
    \SBinRel{R_{\vec{v}}}{\Upsilon}{\vec{E}}{\vec{F}}
  }
  \tag{Congruence}
\]

@@

Now, let us return briefly to Martin-L\"of's notion of type, which we call a
\emph{value type}. The minimal data of such a type's definition consists in a
partial equivalence relation on values,
$\Of{\phi}{\Hom{\SemU}{\Coyoneda{\Upsilon}}{\PERs{\Values_\tau}}}$.  It
is easy to turn such a relation into a computational type
$\IsType{\Upsilon}{\VToCType(\phi)}{\tau}$, as follows, where $\lambda(\chi)$
is the exponential transpose of $\chi$:
\[
  \begin{tikzcd}[sep=huge]
    \IMode{\Coyoneda{\Upsilon}\times\Values_\tau^2}
      \arrow[r, "\IMode{\lambda(\phi)}"]
      \arrow[d, hook, swap, "\IMode{\Pair{\ArrId{}}{i}}"]
      &
    \IMode{\Omega}\\
    \IMode{\Coyoneda{\Upsilon}\times\Exprs_\tau^2}
    \arrow[ur, swap, dotted, "\OMode{\lambda(\VToCType(\phi))}"]
  \end{tikzcd}
\]
where $\VToCType(\phi)$ is the computational extension of $\phi$ in the following sense:
\begin{align*}
  \ADefine{
    \VToCType(\phi)(M,N)
  }{
    \forall M', N' \in \Values_\tau.\
    M\sim_\tau M' \land N\sim_\tau N' \implies \phi(M',N')
  }
\end{align*}
